Enhance Road Detection Data Processing of LiDAR Point Clouds to Specifically Identify Unmarked Gravel Rural Roads

### 2023-05-03 Huston_Paper_asme.tex
methodolgy draft complete???

\hypersetup{%
pdfauthor={John H. Lienhard},
pdftitle={ASME Conference Paper LaTeX Template},                  % <=== change to YOUR pdf file title
pdfkeywords={ASME conference paper, LaTeX template, BibTeX style},% <=== change to YOUR pdf keywords
pdfsubject = {Describes the asmeconf LaTeX template},			  % <=== change to YOUR subject
%	pdfurl={https://ctan.org/pkg/asmeconf},% may delete
pdflicenseurl={https://ctan.org/pkg/asmeconf},% may delete
}

@@

		{Followed methodology to accomplish this work is outlined in Figure \ref{fig:flowz_2}. The experimental apperatus for this work is the Ohio University Autonomous Van van (Figure \ref{fig:van}), which has a Velodyne VLP-32C scanning LiDAR sensor, a Novatel PwrPak 7D-E2 GNSS and INS system, and Mako cameras. Raw data was gathered using the Robotic Operating System (ROS) to package incoming scanning LiDAR, GNSS, IMU, and camera sensor data as rosbags. Scanning LiDAR data was sent from the sensor as a single 360 degree scan at 20 Hz. GNSS and IMU data was captured at a rate of 100 Hz. Rosbags were unpacked using MATLAB tools to separate data streams. Prior to training data extraction, manual dictation of gravel and asphalt areas is necessary. In order to do so, all scanning LiDAR data from a rosbag was compiled in order to create a single point cloud map. Scanning LiDAR timestamps were matched to the closest GNSS and IMU timestamp data. Closest matching GNSS and IMU data informed the location and orientation of the LiDAR scan (Figure \ref{fig:Combined_Classified_PCD}). Compiled point cloud maps may then be examined and compared to camera and satellite data to determine gravel and asphalt surface location. Manually defined 2D areas representing gravel and asphalt were projected unto the point cloud (Figure \ref{fig:Classified_Areas_PCD}).}

		{A single fixed arc from three scanning LiDAR channels in front of the experimental apparatus were considered for classification. Future work may incorporate additional channels. Training data was extracted by examining each 360 scan and comparing the location of the three arcs and the manually projected 2D areas. If the arc resides within the manually defined surface area, raw spatial and intensity values from the arc were saved to a raw training database. Each channel had a separate training data base. Features that describe the raw spatial and intensity data were extracted by performing mathematical functions and saving the results to a feature data base. Thirty percent of the training data base was randomly selected and set aside as a validation data base. Training data was visually examined for clustering to verify that a Random Decision Forest would be able to be trained (Figure \ref{fig:training_data_cluster}). It was found that each class generally clustered to a certain range with some overlap, which indicates that a Random Decision Forest may be trained. MATLAB was used to train a Random Decision Forests for each of the three considered channels using Bayesian Optimization to tune tree depth and number of splits. Training continued until a minimum error was reached (Figure \ref{fig:c2_min_class_error}).}

		{Validation of the }

@@

	\subsection{Detailed Methodology}

@@

@@

@@

	{Scanning LiDAR training data was extracted from manually defined gravel, asphalt, and unknown terrain areas on compiled point clouds. Compiling LiDAR data into a single point cloud was completed by post processing rosbags that contained scanning LiDAR, GPS, and IMU data of an unmarked road. Point cloud translation and rotation was accomplished using transformation matrices derived from GPS and IMU data. Rotation matrices were created using extracted roll, pitch, and yaw data from the IMU. LiDAR origin was found using GPS longitude, latitude, and altitude data. Physical distances between the GPS, IMU, and LiDAR was rectitude by obtaining reference frames provided by AutonomouStuff. GPS coordinates were offset by the current orientation and converted the ground truth to the LiDAR frame. GPS, IMU, and LiDAR reference frames and rotational updates were combined into trajectory vectors. Consecutive point cloud scans may then translated and rotated with derived transformation matrices (Figure \ref{fig:Compiled_PCD}). While not as robust as more sophisticated methods of point cloud aggregation such as NDT or ICP scan matching, over shorter distances using this method proved to be adequate for scoring purposes. Manually defined areas may then be overlayed unto the point cloud (Figure \ref{fig:Combined_Classified_PCD}). Six degree arcs from the first three channels of scanning LiDAR data that lay directly in front of the vehicle were extracted and exported to a database if coincident with the manually defined areas.}

@@

	{MATLAB's Classification Learner Application was used to generate a Random Decision Forests for each considered scanning LiDAR channel. Decision tree depth, number of learners, and number of features to consider for each binary split are hyper-parameters that were optimized using MATLAB's Bayesian Optimization.}

	{Classifying consecutive LiDAR scans was accomplished by examination of nine arcs of interest in front and forty-five degree angles left and right from the front of the experimental apparatus. Features from each arc were extracted and the arc was classified. Averaged x, y, and z coordinates were taken of each arc and projected unto a manually defined truth area map (Figure \ref{fig:Classified_Areas_PCD}). Accuracy scores of each area was calculated, allowing evaluation of road detection performance.}

>>>>

\hypersetup{%
pdfauthor={Rhett Huston},
pdftitle={LiDAR Stuff},                  % <=== change to YOUR pdf file title
}

@@

	{Followed methodology to accomplish this work is outlined in Figure \ref{fig:flowz_2}. The experimental apparatus for this work was the Ohio University Autonomous Van van (Figure \ref{fig:van}), which has a Velodyne VLP-32C scanning LiDAR sensor, a Novatel PwrPak 7D-E2 GNSS and INS system, and Mako cameras. Raw data was gathered using the Robotic Operating System (ROS) to package incoming scanning LiDAR, GNSS, IMU, and camera sensor data as rosbags. Scanning LiDAR data was sent from the sensor as a single 360 degree scan at 20 Hz. GNSS and IMU data was captured at a rate of 100 Hz. Rosbags were unpacked using MATLAB tools to separate data streams. Prior to training data extraction, manual dictation of gravel and asphalt areas is necessary. In order to do so, all scanning LiDAR data from a rosbag was compiled in order to create a single point cloud map. Scanning LiDAR timestamps were matched to the closest GNSS and IMU timestamp data. Closest matching GNSS and IMU data informed the location and orientation of the LiDAR scan (Figure \ref{fig:Combined_Classified_PCD}). Compiled point cloud maps may then be examined and compared to camera and satellite data to determine gravel and asphalt surface location. Manually defined 2D areas representing gravel and asphalt were projected unto the point cloud (Figure \ref{fig:Classified_Areas_PCD}).}

	{Training data was extracted by examining each 360 scan and comparing the location of a 6 degree arc directly in front of the area and the manually projected 2D areas. If the arc resides within the manually defined surface area, raw spatial and intensity values from the arc were saved to a raw training database. Seperate training data bases were used for each individual channel. Features that describe the raw spatial and intensity data were extracted by performing mathematical functions and saving the results to a feature data base. Thirty percent of the training data base was randomly selected and set aside as a validation data base. Training data was visually examined for clustering to verify that a Random Decision Forest would be able to be trained (Figure \ref{fig:training_data_cluster}). It was found that each class generally clustered to a certain range with some overlap, which indicates that a Random Decision Forest may be trained. MATLAB was used to train a Random Decision Forests for each of the three considered channels using Bayesian Optimization to tune tree depth and number of splits. Training continued until a minimum error was reached (Figure \ref{fig:c2_min_class_error}). Validation of the RDF was then completed using the validation database (Figure \ref{fig:vali_conf_mat}). }

	{Classification of five rosbags with data gathered from a section of road with three intercepting gravel driveways (one of the gravel driveways was the entrance to the source of the gravel training data). Three arcs from three channels were considered for classification - the centers of the arcs were directly in front of the vehicle, and 45 degrees left and right of center, creating nine total areas in a 3x3 grid in front of the van. Following the same process as aggregating point clouds, GNSS and IMU data informed the location and orientation of the arcs in each 360 degree scan.}

	{Manually defined gravel, asphalt, and side-of-road areas were projected onto the averaged coordinates of all the classified arcs. Manually defined gravel, asphalt, and side-of-road areas were then examined for the number of type of classified arcs, yielding the accuracy scores for each area. Classified point clouds were then manually examined to determine if the location of gravel and asphalt surfaces may be identified. If consecutive points from each channel in any direction had two or more matching gravel or asphalt classification and was evenly spaced, a gravel or road surface area was projected. Guessed road surface areas were then compared to actual areas to determine the method's ability at unmarked gravel road discovery.}

@@

@@

	\subsection{Data Collection}

@@

	\subsection{Training Data Extraction}

@@

	{Scanning LiDAR training data was extracted from manually defined gravel, asphalt, and unknown terrain areas on compiled point clouds. Compiling LiDAR data into a single point cloud was completed by post processing rosbags that contained scanning LiDAR, GPS, and IMU data of an unmarked road. Scanning LiDAR timestamps were matched to the closest GNSS and IMU timestamp data. Closest matching GNSS and IMU data informed the location and orientation of the LiDAR scan. Point cloud translation and rotation was accomplished using transformation matrices derived from GPS and IMU data. Rotation matrices were created using extracted roll, pitch, and yaw data from the IMU. LiDAR origin was found using GPS longitude, latitude, and altitude data. Physical distances between the GPS, IMU, and LiDAR sensors was rectified by obtaining reference frames provided by AutonomouStuff. GPS coordinates were offset by the current orientation and converted the ground truth to the LiDAR frame. GPS, IMU, and LiDAR reference frames and rotational updates were combined into trajectory vectors. Consecutive point cloud scans may then translated and rotated with derived transformation matrices (Figure \ref{fig:Compiled_PCD}). While not as robust as more sophisticated methods of point cloud aggregation such as NDT or ICP scan matching, over shorter distances using this method proved to be adequate for this work. Manually defined areas may then be overlayed unto the point cloud (Figure \ref{fig:Combined_Classified_PCD}). Six degree arcs from the first three channels of scanning LiDAR data that lay directly in front of the vehicle were extracted and exported to a database if coincident with the manually defined areas.}

@@

	\subsection{RDF Training and Testing}

	{MATLAB's Classification Learner Application was used to generate a Random Decision Forests for each considered scanning LiDAR channel. Decision tree depth, number of learners, and number of features to consider for each binary split are hyper-parameters that were optimized using MATLAB's Bayesian Optimization. All arcs in a channel used the same RDF. }

	{Classifying consecutive LiDAR scans was accomplished by examination of the nine arcs of interest in front and forty-five degree angles left and right from the front of the experimental apparatus. Features from each arc were extracted and the arc was classified. Averaged x, y, and z coordinates were taken of each arc and projected unto a manually defined truth area map (Figure \ref{fig:Classified_Areas_PCD}). Accuracy scores of each area was calculated, allowing evaluation of road detection performance.}
